Learning Discrete-valued Bayesian Networks from Mixed Data

### 2015-11-20 BN_discretizer_paper.tex
fixed Alg 1 and commented out heuristic

\begin{algorithm}
  \caption{Discretization of one continuous variable in a Bayesian network}
  \label{alg:disc_one}
  \begin{algorithmic}[5]
    \Function{DiscretizeOne}{$\cX$, $D$, $G$}
      \State $n \leftarrow$ the number of data instances
      \State $m \leftarrow$ the number of unique values in $D_{\cX}$
      \State $H \leftarrow$ an $N \times N$ matrix such that $H[u,v] = h(u,v)$ (see Algorithm~\ref{alg:h} in Appendix)
      \State $s_{1:m} \leftarrow$ an increasing sequence of integers defined in Section~\ref{subsec:disc_policy}
      \State $L \leftarrow$ the largest cardinality over all discrete variables in the Markov blanket of $X$
      \State $S[u] \leftarrow$ the optimal objective value of subproblems from instance $1$ to $s_u$
      \State $\Lambda[u] \leftarrow$ the optimal discretization for subproblems from instance $1$ to $s_u$
      \State $W[u]  \leftarrow - \ln\brock{1 - {\exp\paren{- L \cdot{ {{\cx_{s_u+1} - \cx_{s_u}}\over{\cx_n - \cx_1}}}}}}$ for $u = 1,2, \ldots,m-1$ and $W[n] \leftarrow 0$
      \For {$v \leftarrow 1$ to $m$}
        \If {$v = 1$}
          \State $S[v] \leftarrow H \paren{1,s_v} + L[v]$
          \State $\Lambda[v] \leftarrow \{ ({\cx_{s_v} + \cx_{s_v+1}}) / 2\}$
        \Else
          \State $s \leftarrow \infty$
          \State $\text{DiscEdge} \leftarrow \infty$
        \For {$u \leftarrow 1$ to $v$}
        \State $s' \leftarrow S[u] + H \paren{s_u+1, s_v} +  {L \cdot {\frac{\cx_{s_v} - \cx_{s_u + 1}}{\cx_n - \cx_1}}} + W[v]$
        \If {$s' < s$}
          \State $s \leftarrow s'$
          \State $\text{DiscEdge} \leftarrow ({x_{s_u} + x_{s_u+1}}) / 2$
        \EndIf
        \EndFor
          \State $S[v] \leftarrow s$
          \State $\Lambda[v] \leftarrow \Lambda[u] \cup \{ DiscEdge \}$
        \EndIf
      \EndFor
      \State \Return $\Lambda[m]$
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\subsection{Approximation}

Algorithm~\ref{alg:disc_one} has an exponential $\bigo\paren{\paren{L'}^{n_p} \cdot N^2}$ runtime which severely limits discretization in networks with a large number of discretization intervals or parent variables.
This section introduces an approximation to the objective function that significantly reduces the runtime and still preserves the quality of discretization.
The approximation replaces the dominator of the last factor in Equation~\ref{eq:likelihood_one} with

\begin{equation}
{\frac{{\gamma_i}!}{ {n^{(P)}_{i,1} !} {n^{(P)}_{i,2} !} \cdots {n^{(P)}_{i,J_P} !}}} \approx \prod_{r=1}^{n_p} { {{{\gamma}_i}!}\over{ {n^{(P_r)}_{i,1} !} {n^{(P_r)}_{i,2} !} \cdots {n^{(P_r)}_{i,J_{P_r}} !}}}\text{,}
\end{equation}

\noindent
where $P_r$ is the $r$th parent variable, $J_{p_r} = \| P_r\|$, and ${n^{(P_r)}_{i,j} !}$ is the number of instances in the $i$th interval with the $j$th value of $P_r$. Note that $\gamma_i = \sum_j {n^{(P_r)}_{i,j}}$ for all $r$.
The approximated objective function is

\begin{small}
  \begin{equation}
  \label{eq:opt_prob_approx}
  \begin{aligned}
    & \sum_{i=1}^{k-1}
     - \ln
      \paren{
        1 - \exp
        \paren{
          - L \cdot  \frac{
                           \cx_{\lambda_i +1} - \cx_{\lambda_i}
                         }{
                           \cx_n - \cx_1
                          }
        }
      }
        + \sum_{i=1}^{k}
        L \cdot \frac{\cx_{\lambda_{i}} - \cx_{\lambda_{i-1} + 1}}{\cx_n - \cx_1} \\
    & \sum_{i=1}^{k} \brock{
    \ln{{\gamma_i + J_P - 1}\choose{J_P - 1}}
    +{ \ln \paren{
          \frac{
            {\gamma_i}!
          }{ \sum_{r=1}^{n_p}
            {n^{({P_r})}_{i,1}}! \; {n^{({P_r})}_{i,2}}! \; \cdots \; {n^{({P_r})}_{i,J_{P_r}}}!
          }
        }}
    } \\
    & \sum_{j=1}^{n_c} \sum_{i=1}^k \sum_{l=1}^{J^{(j)}_S} \brock{
    { \ln
            {{n^{(j)}_{i,l} + J_{C_j} - 1}\choose{J_{C_j}-1}}}
      +
      \ln \paren{ {
            \frac{
              {n^{(j)}_{i,l}}!
            }{
              {n^{(j)}_{i,1,l}!} \; {n^{(j)}_{i,2,l}!} \; \cdots \; {n^{(j)}_{i,J_{C_j},l}!}
            }
          }
          } }
  \end{aligned}
  \end{equation}
\end{small}

The approximation reduces the runtime of Algorithm~\ref{alg:disc_one} to a polynomial, $\todo{\bigo\paren{ {v'_\text{max}} \cdot (n_c + n_p) \cdot n^2}}$.
It will be shown in Section~\ref{sec:experiments} that the approximation is more sensitive to the distribution over other variables and is biased towards slightly higher interval counts.
\tim{We need to have a more specific thing to say along the lines of `When does it fail?'}

The approximated objective function, Equation~\ref{eq:opt_prob_approx}, provides intuition into how child and parent variables contribute to the objective function.
For example, in the left graph of Figure~\ref{fig:example_networks}, the corresponding square brackets in the approximated objective function are

\begin{small}
  \begin{equation}
  \begin{aligned}
  \label{eq:example_approx}
   \sum_{i=1}^k & \left\lbrace   \left[ \ln{{\gamma_{i} + J_{C_1} - 1}\choose{J_{C_1}-1}} + \ln \left(  {{{\gamma_i}!} \over { {n^{(1)}_{i,1,\emptyset} !} \cdots {n^{(1)}_{i,J_{C_1},\emptyset} !}} }  \right)  +  \right. \right.\\
  & \left.  \ln{{\gamma_{i} + J_{C_2} - 1}\choose{J_{C_2}-1}} + \ln \left(  {{{n_i}!} \over { {n^{(2)}_{i,1,\emptyset} !} \cdots {n^{(2)}_{i,J_{C_1},\emptyset} !}} }  \right)  \right] + \\
  &  \left. \left[  \ln {{\gamma_{i} + J_P - 1}\choose{J_P-1}} +  \ln \left( { {{{\gamma}_i}!}\over{ {n^{(P_1)}_{i,1} !}\cdots {n^{(P_1)}_{i,J_{P_1}} !}}} \right) +\ln \paren{{ {{{\gamma}_i}!}\over{ {n^{(P_2}_{i,1} !}\cdots {n^{(P_2)}_{i,J_{P_2}} !}}} } \right] \right\rbrace \text{.}
  \end{aligned}
  \end{equation}
\end{small}

In Equation~\ref{eq:example_approx}, each child carries two terms, as shown in the first set of square brackets, but the two parent variables only carry a total of three terms, as shown in the second set of square brackets.
In this case the child variables have greater effect on the optimal discretization than the parent variables, even though the number of child and parent variables are the same.
However, if $C_2$ has an additional parent $S_2$, as shown in the right graph of Figure~\ref{fig:example_networks}, the importances of $C_2$ will be reduced.
This is due to the additional influence of $S_2$ on $C_2$.
This phenomenon arrises because the Bayesian method, but with and without the heuristic, incorporates graph structure when discretizing continuous variables.

\begin{figure}[ht]
  \centering
  \begin{tabular}{cc}
    \input{graph2_left}
    \end{tabular}
    \hspace{5em}
    \begin{tabular}{cc}
    \input{graph2_right}
  \end{tabular}
  \caption{Two example networks}
  \label{fig:example_networks}
\end{figure}

@@

\begin{algorithm}
  \caption{Discretization of multiple continuous variables}
  \label{alg:disc_two}
  \begin{algorithmic}[5]
  \Function{DiscretizeAll}{$D$, $G$, $\bm{\cX}$, $\maxpasses$}
    \State $\discset \leftarrow$ the discretization policies for each $\cX$ in $\bm{\cX}$
    \State $D^* \leftarrow $ the dataset $D$ discretized according to $\discset$
    \State $k \leftarrow \textsc{Median}\paren{\curly{ |X| \text{ s.t. } X \text{ does not have corresponding } \cX \text{ in } \bm{\cX} }}$
    \For {$\paren{\cX, X}$ such that $\cX \in \bm{\cX}$ and $X$ is the discretized version of $\cX$}
      \State $\Lambda_{\cX} \leftarrow$ equal-width discretization of $\cX$ with $k$ intervals
      \State $D^*_{X} \leftarrow \Lambda_{\cX} \paren{D_{\cX}}$
    \EndFor
    \State $n_\text{cycle} \leftarrow 0$
    \While {$\discset$ has not converged \textbf{and} $n_\text{cycle} \leq \maxpasses$}
      \State increment {$n_\text{cycle}$}
      \For {$\paren{\cX, X}$ such that $\cX \in \bm{\cX}$}
        \State {$D^*_X \leftarrow D_{\cX}$}
        \State {$\Lambda_{\cX} \leftarrow \textsc{DiscretizeOne}(\cX, D^*, G)$}
        \State {$D^*_X \leftarrow \Lambda_{\cX}\paren{D_{\cX}}$}
      \EndFor
    \EndWhile
    \State \Return $\discset$
  \EndFunction
  \end{algorithmic}
\end{algorithm}

@@

\section{Experiments}
\label{sec:experiments}

This section describes experiments conducted to evaluate the proposed methods.
All experiments were run on datasets from the publically available University of California, Irvine machine learning repository \citep{Lichman_2013}.
Variables are labelled alphabetically in the order given on the dataset information webpage
In the figures that follow, shaded nodes correspond to initially discrete variables and the subscripts indicate the number of discrete instantiations.

Two experiments were conducted on each dataset.
The first experiment compares the performance of the optimal Bayesian discretization method, the heuristic Bayesian discretization method, and MDL on a known Bayesian network structure.
The structure was obtained by prediscretizing each continuous variable into $k$ uniform-width intervals, where $k$ is the median number of instantiations of the discrete variables, and using the structure with the highest Bayesian score from one thousand runs of the K2 algorithm with random variable orderings.
The second experiment compares the same methods applied when simultaneously discretizing and learning network structure.

The discretizations are compared using the mean cross validated log-likelihood of the data $D$ given the graph structure $G$ and discretization policies $\discset$.
The log-likelihood has two components,

\begin{equation}
\ln p(D\mid G, \discset) = \ln P(D^*\mid G) + \ln p(D\mid \discset, D^*)\text{,}
\end{equation}

\noindent
where $D$ is the original dataset and $D^*$ is the dataset discretized according to $\discset$.
The log-likelihood of the discretized dataset is the Bayesian score (Equation~\ref{eq:BayesianScore}) evaluated with a uniform prior over graph structures and a uniform Dirichlet prior ($\alpha_{ijk} = 1$ for all $i$, $j$, and $k$).
The log-likelihood of the original dataset given the discrete dataset is

\todo{
\begin{equation}
  \ln p\paren{D\mid \discset, D^*} = \sum_{i=1}^n
  1\curly{X_i \text{ orig. cont.}}
  \sum_{j=1}^{q_i} \sum_{k=1}^{r_i} m_{ijk} \ln\paren{
    \frac{
      1
    }{
      e^{\Lambda_{\cX_i}}_{k+1} - e^{\Lambda_{\cX_i}}_{k}
    }
  }\text{.}
\end{equation}
}

\noindent
The mean cross-validated log-likelihood is the mean log-likelihood on the witheld dataset among cross-validation folds.
Cross-validation directly estimates generalization error.
Ten folds were used in each experiment.

The optimal discretization policy for MDL is found with dynamic programming in a similar manner as with the Bayesian method.
For a Bayesian network with a single continuous variable $\cX$, the MDL objective function is

\begin{equation}
  \label{eqn:MDL}
  \todo{
  \begin{aligned}
  \frac{1}{2} \ln(n) \left\lbrace  \lvert \Pa_{X_i} \rvert (\lvert X_i \rvert - 1) +
   {\sum_{j,\cX_i \in \Pa_{X_i}}} \lvert \Pa_{X_j} \rvert (\lvert \cX_j \rvert - 1) \right\rbrace + \ln(\lvert {\cX_i} \rvert) \\
   + (N_i-1) H \paren{\frac{|X^*_i| - 1}{N_i -1}} -n \cdot \left[ I(X_i,\Pa_{X_i}) + {\sum_{j,\cX_i \in \Pa_{\cX_i}}} I(\cX_j, \Pa_{X_j}) \right]\text{,}
  \end{aligned}
  }
\end{equation}

where $I(A,B)$ is the mutual information between two discrete variables $A$ and $B$, ${H(p) = -p \ln(p) - (1-p) \ln(1-p)}$, and \todo{$X^*_i$ is the discretized version of the $i$th sample of $X$}.

\subsection{Auto MPG Dataset}
\label{subsec:auto}

The Auto MPG dataset contains variables related to the fuel consumption of automobiles in urban driving.
The dataset has \num{392} samples over eight variables, not including six instances with missing data.
Three variables are discrete: $B$, $G$, and $H$, with \num{5}, \num{13}, and \num{3} instantiations respectively.

\subsubsection{Discretization with fixed Structure}
\label{subsubsec:auto_exp1}

The Bayesian approaches and the MDL discretization method were tested on the Auto MPG data using the network network shown in Figure~\ref{fig:auto_graph_1}.
This structure was obtained by prediscretizing each continuous variable into five uniform-width intervals, where five is the median cardinality of the discrete variables, and then taking the structure with highest likelihood from one thousand runs of K2.

\begin{figure}[ht]
  \centering
   \input{graph_autocar_1}
   \caption{Bayesian network structure obtained from running K2 on the prediscretized Auto MPG dataset}
  \label{fig:auto_graph_1}
\end{figure}

Table~\ref{table:auto_disc_table_1} lists the discretization edges and mean log-likelihoods under 10-fold cross validation of the discrete Bayesian network resulting from each discretization method.
The results indicate that the heuristic Bayesian method produces exactly the same discretization policy as the optimal Bayesian method.
The MDL method does not produce any discretization edges, assigning one continuous interval to each continuous variable, and produces the structure with the lowest score.
The cause behind MDL's discretization policy is discussed in Section~\ref{subsec:discuss_exp}.

\begin{table}[h]
  \centering
  \caption{
    Results from discretization of the Auto MPG dataset with fixed structure from Figure~\ref{fig:auto_graph_1}.
    The first five rows list the discretization edges for each continuous variable.
    The last row lists the mean cross-validated log-likelihood; positive values are better.
    The optimal and heuristic Bayesian discretization policies are identical and the MDL method does not produce any discretization edges.
  }
  \input{table_autoMPG_1.tex}
  \label{table:auto_disc_table_1}
\end{table}

Figure~\ref{fig:auto_exp1_distr_1_3} compares the Bayesian discretizaton policy for variables $A$ and $C$ with the original Auto MPG data.
The black marks are the original Auto MPG samples.
The color of a discretized region indicates the marginal probability of a sample from $P(A,C)$ being drawn from that region.
As shown in the figure, the discretization edges obtained by the optimal Bayesian approach and the resulting marginalized probability match the raw data well.

\begin{figure}[ht]
\centering
  \input{graph_auto_exp1_distr_1_3}
  \caption{
    Comparison of the Bayesian discretization policy for variables $A$ and $C$ to the original Auto MPG data learned with a fixed network.
    The marginal probabilities of discretized regions closely match the raw data.
  }
  \label{fig:auto_exp1_distr_1_3}
\end{figure}

\subsubsection{Discretization while Learning Structure}
\label{subsubsec:auto_exp2}

In this experiment the network structure was not fixed in advance and was learned simultaneously with the discretization policies.
Figure~\ref{fig:auto_graph_2} shows a learned Bayesian network structure and the corresponding numbers of intervals after discretization for each continuous variable.
This result was obtained by running Algorithm~\ref{alg:structure_learn} fifty times using the optimal Bayesian method and choosing the structure with the highest likelihood.

\begin{figure}[ht]
  \centering
  \input{graph_autocar_2}
  \caption{The discrete-valued Bayesian network learned from the Auto MPG dataset}
  \label{fig:auto_graph_2}
\end{figure}

Figure~\ref{fig:auto_exp2_distr_1_3} compares the Bayesian discretizaton policy for variables $A$ and $C$ in the learned network with the original Auto MPG data.
The color of a discretized region indicates the marginal probability of a sample from $P(A,C)$ being drawn from that region.
Although there are fewer discretization edges for $A$ and $C$ in the learned network, the marginal distribution is still captured.
Although the discretization result of the optimal Bayesian approach may vary according to the network structure, it still produces high-quality discretizations.

\begin{figure}[ht]
  \centering
  \input{graph_auto_exp2_distr_1_3}
  \caption{
    Comparison of the optimal Bayesian discretization policy for variables $A$ and $C$ to the original Auto MPG data learned along with the network structure.
    Although the number of discretization edges is less than those in Figure~\ref{fig:auto_exp1_distr_1_3}, the probability distribution still closely matches the original data.
  }
  \label{fig:auto_exp2_distr_1_3}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% WINE
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\begin{algorithm}
  \caption{Discretization of one continuous variable in a Bayesian network}
  \label{alg:disc_one}
  \begin{algorithmic}[5]
    \Function{DiscretizeOne}{$D$, $G$, $\cX$}
      % \State $n \leftarrow$ the number of data instances
      % \State $m \leftarrow$ the number of unique values in $D_{\cX}$
      \State $H \leftarrow$ an $n \times n$ matrix such that $H[u,v] = h(u,v)$; can be precomputed
      % \State $s_{1:m} \leftarrow$ an increasing sequence of integers defined in Section~\ref{subsec:disc_policy}; can be precomputed
      \State $L \leftarrow$ the largest cardinality over all discrete variables in the Markov blanket of $\cX$
      \State $S[i] \leftarrow$ the optimal objective value computed over samples $1$ to $s_i$
      \State $\Lambda[i] \leftarrow$ the discretization policy for the subproblem over samples $1$ to $s_i$
      \State $W[i]  \leftarrow - \ln\brock{1 - {\exp\paren{- L \cdot{ {{\cx_{s_i+1} - \cx_{s_i}}\over{\cu_m - \cu_1}}}}}}$ for $i \in [1,m-1]$ and $W[m] \leftarrow 0$
      \For {$v \leftarrow 1$ to $m$}
        \If {$v = 1$}
          \State $S[v] \leftarrow H \paren{1,s_v} + L[v]$
          \State $\Lambda[v] \leftarrow \curly{({\cu_v + \cu_{v+1}}) / 2}$
        \Else
          \State $\hat{S}, \hat{u} \leftarrow \infty, 0$
          \State $\text{DiscEdge} \leftarrow \infty$
          \For {$u \leftarrow 1$ to $v$}
            \If {$u = v$}
              \State $\tilde{S} \leftarrow W[v] + H \paren{1, s_v} +  {L \cdot {\frac{\cu_{v} - \cu_1}{\cu_m - \cu_1}}}$
            \Else
              \State $\tilde{S} \leftarrow W[v] + H \paren{s_u+1, s_v} +  {L \cdot {\frac{\cu_{v} - \cu_{u + 1}}{\cu_m - \cu_1}}} + S[u]$
            \EndIf
            \If {$\tilde{S} < \hat{S}$}
              \State $\hat{S}, \hat{u} \leftarrow \tilde{S}, u$
              \State $\text{DiscEdge} \leftarrow ({x_{s_u} + x_{s_u+1}}) / 2$
            \EndIf
          \EndFor
          \State $S[v] \leftarrow \hat{S}$
          \State $\Lambda[v] \leftarrow \Lambda[u] \cup \{ \text{DiscEdge} \}$
        \EndIf
      \EndFor
      \State \Return $\Lambda[m]$
    \EndFunction
  \end{algorithmic}
\end{algorithm}

% \subsection{Approximation}

% Algorithm~\ref{alg:disc_one} has an exponential $\bigo\paren{\paren{L'}^{n_p} \cdot N^2}$ runtime which severely limits discretization in networks with a large number of discretization intervals or parent variables.
% This section introduces an approximation to the objective function that significantly reduces the runtime and still preserves the quality of discretization.
% The approximation replaces the dominator of the last factor in Equation~\ref{eq:likelihood_one} with

% \begin{equation}
% {\frac{{\gamma_i}!}{ {n^{(P)}_{i,1} !} {n^{(P)}_{i,2} !} \cdots {n^{(P)}_{i,J_P} !}}} \approx \prod_{r=1}^{n_p} { {{{\gamma}_i}!}\over{ {n^{(P_r)}_{i,1} !} {n^{(P_r)}_{i,2} !} \cdots {n^{(P_r)}_{i,J_{P_r}} !}}}\text{,}
% \end{equation}

% \noindent
% where $P_r$ is the $r$th parent variable, $J_{p_r} = \| P_r\|$, and ${n^{(P_r)}_{i,j} !}$ is the number of instances in the $i$th interval with the $j$th value of $P_r$. Note that $\gamma_i = \sum_j {n^{(P_r)}_{i,j}}$ for all $r$.
% The approximated objective function is

% \begin{small}
%   \begin{equation}
%   \label{eq:opt_prob_approx}
%   \begin{aligned}
%     & \sum_{i=1}^{k-1}
%      - \ln
%       \paren{
%         1 - \exp
%         \paren{
%           - L \cdot  \frac{
%                            \cx_{\lambda_i +1} - \cx_{\lambda_i}
%                          }{
%                            \cx_n - \cx_1
%                           }
%         }
%       }
%         + \sum_{i=1}^{k}
%         L \cdot \frac{\cx_{\lambda_{i}} - \cx_{\lambda_{i-1} + 1}}{\cx_n - \cx_1} \\
%     & \sum_{i=1}^{k} \brock{
%     \ln{{\gamma_i + J_P - 1}\choose{J_P - 1}}
%     +{ \ln \paren{
%           \frac{
%             {\gamma_i}!
%           }{ \sum_{r=1}^{n_p}
%             {n^{({P_r})}_{i,1}}! \; {n^{({P_r})}_{i,2}}! \; \cdots \; {n^{({P_r})}_{i,J_{P_r}}}!
%           }
%         }}
%     } \\
%     & \sum_{j=1}^{n_c} \sum_{i=1}^k \sum_{l=1}^{J^{(j)}_S} \brock{
%     { \ln
%             {{n^{(j)}_{i,l} + J_{C_j} - 1}\choose{J_{C_j}-1}}}
%       +
%       \ln \paren{ {
%             \frac{
%               {n^{(j)}_{i,l}}!
%             }{
%               {n^{(j)}_{i,1,l}!} \; {n^{(j)}_{i,2,l}!} \; \cdots \; {n^{(j)}_{i,J_{C_j},l}!}
%             }
%           }
%           } }
%   \end{aligned}
%   \end{equation}
% \end{small}

% The approximation reduces the runtime of Algorithm~\ref{alg:disc_one} to a polynomial, $\todo{\bigo\paren{ {v'_\text{max}} \cdot (n_c + n_p) \cdot n^2}}$.
% It will be shown in Section~\ref{sec:experiments} that the approximation is more sensitive to the distribution over other variables and is biased towards slightly higher interval counts.
% \tim{We need to have a more specific thing to say along the lines of `When does it fail?'}

% The approximated objective function, Equation~\ref{eq:opt_prob_approx}, provides intuition into how child and parent variables contribute to the objective function.
% For example, in the left graph of Figure~\ref{fig:example_networks}, the corresponding square brackets in the approximated objective function are

% \begin{small}
%   \begin{equation}
%   \begin{aligned}
%   \label{eq:example_approx}
%    \sum_{i=1}^k & \left\lbrace   \left[ \ln{{\gamma_{i} + J_{C_1} - 1}\choose{J_{C_1}-1}} + \ln \left(  {{{\gamma_i}!} \over { {n^{(1)}_{i,1,\emptyset} !} \cdots {n^{(1)}_{i,J_{C_1},\emptyset} !}} }  \right)  +  \right. \right.\\
%   & \left.  \ln{{\gamma_{i} + J_{C_2} - 1}\choose{J_{C_2}-1}} + \ln \left(  {{{n_i}!} \over { {n^{(2)}_{i,1,\emptyset} !} \cdots {n^{(2)}_{i,J_{C_1},\emptyset} !}} }  \right)  \right] + \\
%   &  \left. \left[  \ln {{\gamma_{i} + J_P - 1}\choose{J_P-1}} +  \ln \left( { {{{\gamma}_i}!}\over{ {n^{(P_1)}_{i,1} !}\cdots {n^{(P_1)}_{i,J_{P_1}} !}}} \right) +\ln \paren{{ {{{\gamma}_i}!}\over{ {n^{(P_2}_{i,1} !}\cdots {n^{(P_2)}_{i,J_{P_2}} !}}} } \right] \right\rbrace \text{.}
%   \end{aligned}
%   \end{equation}
% \end{small}

% In Equation~\ref{eq:example_approx}, each child carries two terms, as shown in the first set of square brackets, but the two parent variables only carry a total of three terms, as shown in the second set of square brackets.
% In this case the child variables have greater effect on the optimal discretization than the parent variables, even though the number of child and parent variables are the same.
% However, if $C_2$ has an additional parent $S_2$, as shown in the right graph of Figure~\ref{fig:example_networks}, the importances of $C_2$ will be reduced.
% This is due to the additional influence of $S_2$ on $C_2$.
% This phenomenon arrises because the Bayesian method, but with and without the heuristic, incorporates graph structure when discretizing continuous variables.

% \begin{figure}[ht]
%   \centering
%   \begin{tabular}{cc}
%     \input{graph2_left}
%     \end{tabular}
%     \hspace{5em}
%     \begin{tabular}{cc}
%     \input{graph2_right}
%   \end{tabular}
%   \caption{Two example networks}
%   \label{fig:example_networks}
% \end{figure}

@@

\begin{algorithm}
  \caption{Discretization of multiple continuous variables}
  \label{alg:disc_two}
  \begin{algorithmic}[5]
  \Function{DiscretizeAll}{$D$, $G$, $\bm{\cX}$, $\maxpasses$}
    \State $\discset \leftarrow$ the discretization policies for each $\cX$ in $\bm{\cX}$
    \State $D^* \leftarrow $ the dataset $D$ discretized according to $\discset$
    \State $k \leftarrow \textsc{Median}\paren{\curly{ |X| \text{ s.t. } X \text{ does not have corresponding } \cX \text{ in } \bm{\cX} }}$
    \For {$\paren{\cX, X}$ such that $\cX \in \bm{\cX}$ and $X$ is the discretized version of $\cX$}
      \State $\Lambda_{\cX} \leftarrow$ equal-width discretization of $\cX$ with $k$ intervals
      \State $D^*_{X} \leftarrow \Lambda_{\cX} \paren{D_{\cX}}$
    \EndFor
    \State $n_\text{cycle} \leftarrow 0$
    \While {$\discset$ has not converged \textbf{and} $n_\text{cycle} \leq \maxpasses$}
      \State increment {$n_\text{cycle}$}
      \For {$\paren{\cX, X}$ such that $\cX \in \bm{\cX}$}
        \State {$D^*_X \leftarrow D_{\cX}$}
        \State {$\Lambda_{\cX} \leftarrow \textsc{DiscretizeOne}(D^*, G, \cX)$}
        \State {$D^*_X \leftarrow \Lambda_{\cX}\paren{D_{\cX}}$}
      \EndFor
    \EndWhile
    \State \Return $\discset$
  \EndFunction
  \end{algorithmic}
\end{algorithm}

@@

% \section{Experiments}
% \label{sec:experiments}

% This section describes experiments conducted to evaluate the proposed methods.
% All experiments were run on datasets from the publically available University of California, Irvine machine learning repository \citep{Lichman_2013}.
% Variables are labelled alphabetically in the order given on the dataset information webpage
% In the figures that follow, shaded nodes correspond to initially discrete variables and the subscripts indicate the number of discrete instantiations.

% Two experiments were conducted on each dataset.
% The first experiment compares the performance of the optimal Bayesian discretization method, the heuristic Bayesian discretization method, and MDL on a known Bayesian network structure.
% The structure was obtained by prediscretizing each continuous variable into $k$ uniform-width intervals, where $k$ is the median number of instantiations of the discrete variables, and using the structure with the highest Bayesian score from one thousand runs of the K2 algorithm with random variable orderings.
% The second experiment compares the same methods applied when simultaneously discretizing and learning network structure.

% The discretizations are compared using the mean cross validated log-likelihood of the data $D$ given the graph structure $G$ and discretization policies $\discset$.
% The log-likelihood has two components,

% \begin{equation}
% \ln p(D\mid G, \discset) = \ln P(D^*\mid G) + \ln p(D\mid \discset, D^*)\text{,}
% \end{equation}

% \noindent
% where $D$ is the original dataset and $D^*$ is the dataset discretized according to $\discset$.
% The log-likelihood of the discretized dataset is the Bayesian score (Equation~\ref{eq:BayesianScore}) evaluated with a uniform prior over graph structures and a uniform Dirichlet prior ($\alpha_{ijk} = 1$ for all $i$, $j$, and $k$).
% The log-likelihood of the original dataset given the discrete dataset is

% \todo{
% \begin{equation}
%   \ln p\paren{D\mid \discset, D^*} = \sum_{i=1}^n
%   1\curly{X_i \text{ orig. cont.}}
%   \sum_{j=1}^{q_i} \sum_{k=1}^{r_i} m_{ijk} \ln\paren{
%     \frac{
%       1
%     }{
%       e^{\Lambda_{\cX_i}}_{k+1} - e^{\Lambda_{\cX_i}}_{k}
%     }
%   }\text{.}
% \end{equation}
% }

% \noindent
% The mean cross-validated log-likelihood is the mean log-likelihood on the witheld dataset among cross-validation folds.
% Cross-validation directly estimates generalization error.
% Ten folds were used in each experiment.

% The optimal discretization policy for MDL is found with dynamic programming in a similar manner as with the Bayesian method.
% For a Bayesian network with a single continuous variable $\cX$, the MDL objective function is

% \begin{equation}
%   \label{eqn:MDL}
%   \todo{
%   \begin{aligned}
%   \frac{1}{2} \ln(n) \left\lbrace  \lvert \Pa_{X_i} \rvert (\lvert X_i \rvert - 1) +
%    {\sum_{j,\cX_i \in \Pa_{X_i}}} \lvert \Pa_{X_j} \rvert (\lvert \cX_j \rvert - 1) \right\rbrace + \ln(\lvert {\cX_i} \rvert) \\
%    + (N_i-1) H \paren{\frac{|X^*_i| - 1}{N_i -1}} -n \cdot \left[ I(X_i,\Pa_{X_i}) + {\sum_{j,\cX_i \in \Pa_{\cX_i}}} I(\cX_j, \Pa_{X_j}) \right]\text{,}
%   \end{aligned}
%   }
% \end{equation}

% where $I(A,B)$ is the mutual information between two discrete variables $A$ and $B$, ${H(p) = -p \ln(p) - (1-p) \ln(1-p)}$, and \todo{$X^*_i$ is the discretized version of the $i$th sample of $X$}.

% \subsection{Auto MPG Dataset}
% \label{subsec:auto}

% The Auto MPG dataset contains variables related to the fuel consumption of automobiles in urban driving.
% The dataset has \num{392} samples over eight variables, not including six instances with missing data.
% Three variables are discrete: $B$, $G$, and $H$, with \num{5}, \num{13}, and \num{3} instantiations respectively.

% \subsubsection{Discretization with fixed Structure}
% \label{subsubsec:auto_exp1}

% The Bayesian approaches and the MDL discretization method were tested on the Auto MPG data using the network network shown in Figure~\ref{fig:auto_graph_1}.
% This structure was obtained by prediscretizing each continuous variable into five uniform-width intervals, where five is the median cardinality of the discrete variables, and then taking the structure with highest likelihood from one thousand runs of K2.

% \begin{figure}[ht]
%   \centering
%    \input{graph_autocar_1}
%    \caption{Bayesian network structure obtained from running K2 on the prediscretized Auto MPG dataset}
%   \label{fig:auto_graph_1}
% \end{figure}

% Table~\ref{table:auto_disc_table_1} lists the discretization edges and mean log-likelihoods under 10-fold cross validation of the discrete Bayesian network resulting from each discretization method.
% The results indicate that the heuristic Bayesian method produces exactly the same discretization policy as the optimal Bayesian method.
% The MDL method does not produce any discretization edges, assigning one continuous interval to each continuous variable, and produces the structure with the lowest score.
% The cause behind MDL's discretization policy is discussed in Section~\ref{subsec:discuss_exp}.

% \begin{table}[h]
%   \centering
%   \caption{
%     Results from discretization of the Auto MPG dataset with fixed structure from Figure~\ref{fig:auto_graph_1}.
%     The first five rows list the discretization edges for each continuous variable.
%     The last row lists the mean cross-validated log-likelihood; positive values are better.
%     The optimal and heuristic Bayesian discretization policies are identical and the MDL method does not produce any discretization edges.
%   }
%   \input{table_autoMPG_1.tex}
%   \label{table:auto_disc_table_1}
% \end{table}

% Figure~\ref{fig:auto_exp1_distr_1_3} compares the Bayesian discretizaton policy for variables $A$ and $C$ with the original Auto MPG data.
% The black marks are the original Auto MPG samples.
% The color of a discretized region indicates the marginal probability of a sample from $P(A,C)$ being drawn from that region.
% As shown in the figure, the discretization edges obtained by the optimal Bayesian approach and the resulting marginalized probability match the raw data well.

% \begin{figure}[ht]
% \centering
%   \input{graph_auto_exp1_distr_1_3}
%   \caption{
%     Comparison of the Bayesian discretization policy for variables $A$ and $C$ to the original Auto MPG data learned with a fixed network.
%     The marginal probabilities of discretized regions closely match the raw data.
%   }
%   \label{fig:auto_exp1_distr_1_3}
% \end{figure}

% \subsubsection{Discretization while Learning Structure}
% \label{subsubsec:auto_exp2}

% In this experiment the network structure was not fixed in advance and was learned simultaneously with the discretization policies.
% Figure~\ref{fig:auto_graph_2} shows a learned Bayesian network structure and the corresponding numbers of intervals after discretization for each continuous variable.
% This result was obtained by running Algorithm~\ref{alg:structure_learn} fifty times using the optimal Bayesian method and choosing the structure with the highest likelihood.

% \begin{figure}[ht]
%   \centering
%   \input{graph_autocar_2}
%   \caption{The discrete-valued Bayesian network learned from the Auto MPG dataset}
%   \label{fig:auto_graph_2}
% \end{figure}

% Figure~\ref{fig:auto_exp2_distr_1_3} compares the Bayesian discretizaton policy for variables $A$ and $C$ in the learned network with the original Auto MPG data.
% The color of a discretized region indicates the marginal probability of a sample from $P(A,C)$ being drawn from that region.
% Although there are fewer discretization edges for $A$ and $C$ in the learned network, the marginal distribution is still captured.
% Although the discretization result of the optimal Bayesian approach may vary according to the network structure, it still produces high-quality discretizations.

% \begin{figure}[ht]
%   \centering
%   \input{graph_auto_exp2_distr_1_3}
%   \caption{
%     Comparison of the optimal Bayesian discretization policy for variables $A$ and $C$ to the original Auto MPG data learned along with the network structure.
%     Although the number of discretization edges is less than those in Figure~\ref{fig:auto_exp1_distr_1_3}, the probability distribution still closely matches the original data.
%   }
%   \label{fig:auto_exp2_distr_1_3}
% \end{figure}
